Tracking and predicting turn structure during language acquisition

### 2014-09-30 paper/CasillasFrank-Anticip.tex
ref order, experiment > study, age expressions

\usepackage{graphicx}
% \usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{tipa}
\usepackage{float}

@@

%\linenumbers

@@

Spontaneous conversation is a universal context for using and learning language. Like other types of human interaction, it is organized at its core by the roles and goals of its participants. But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk. Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities. Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{stivers2009, dingemanse2013, de-vosInPrep}. How does this ability develop?

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking. Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others. And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.

@@

The majority of work on children's early turn taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, snow1977, jaffe2001}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}. Infants at this age also notice changes to social contingency outside of turn structure. In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze. When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.

@@

We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the study. Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school. All participants were native English speakers. Approximately one-third (N=25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from analyses. None of the adult participants reported fluency in a second language.

@@

All participants (child and adult) completed a five-point calibration routine before the first video started. We used a dancing Elmo for the children's calibration image. During the study, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs). The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean. The presentation order of the non-English videos was shuffled into four lists, which participants were assigned to randomly. The entire experiment, including instructions, took 10--15 minutes.

@@

As in the first study, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.

@@

\subsubsection{Participants}
We recruited 27 undergraduate adults and 129 children between ages 1;0--7;0 to participate in our study. We recruited our child participants from the Children's Discovery Museum in San Jose, California, targeting approximately 20 children for each of the six 1-year age groups (range$=$20--23). All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home. None of the adult participants reported fluency in a second language. We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.

@@

\subsection{Procedure}
We used the same experimental apparatus and procedure as in the first study. Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs). Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio. This study required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation. The entire experiment took less than five minutes.

@@

Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax. We conducted two \textit{post-hoc} \textit{t}-tests for children under 3;0, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions. We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407). This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under 3;0. The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}). So, whereas the early advantage for prosody appeared for both transition types (question and non-question), the later-emerging advantage for lexicosyntax and prosody together was primarily limited to question transitions. This pattern of results may suggest a qualitative shift in the kind of predictions children make while observing a conversation, with linguistic cues (like prosody) playing different roles at different ages.

@@

\section{General Discussion}
\label{sec:gendisc}
Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, snow1977, jaffe2001, hilbrinkInPrep}. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.

@@

Our findings are only partially consistent with prior experimental work on turn taking. De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened. Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at age 3;0 (with switches by younger children never exceeding chance). Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax. Furthermore, our adult control participants patterned similarly to the children: They showed question effects for languages they didn't speak (Experiment 1) and showed the strongest question effects for lexicosyntax and prosody together (Experiment 2). Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equal to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior. This may partially be due to the fact that we controlled syllable duration in addition to intonation, thereby eliminating two of the most salient cues to prosodic structure.

@@

Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under 3;0 make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not. Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions. But our results also differ from theirs in a few important ways.

First, we found that children begin making above-chance anticipatory switches much earlier than 3;0---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2. This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally. Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions. Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.

>>>>

\usepackage{graphicx}
\usepackage{lineno}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{tipa}
\usepackage{float}

@@

\linenumbers

@@

Spontaneous conversation is a universal context for using and learning language. Like other types of human interaction, it is organized at its core by the roles and goals of its participants. But what sets conversation apart is its structure: sequences of interconnected, communicative actions that take place across alternating turns at talk. Sequential, turn-based structures in conversation are strikingly uniform across language communities and linguistic modalities. Turn-taking behaviors are also cross-culturally consistent in their basic features and the details of their implementation \citep{de-vosInPrep, dingemanse2013, stivers2009}. How does this ability develop?

Children participate in sequential coordination with their caregivers starting at three months of age---before they can rely on any linguistic cues in taking turns \citep[see, among others, ][]{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}. Of course, infant turn taking is different from adult turn taking in several ways. Infant turn taking is heavily scaffolded by caregivers, has distinct timing in comparison to adult turn taking, and lacks semantic content \citep{hilbrinkInPrep, jaffe2001}. However, children's early, turn-structured social interactions are presumably a critical precursor to their conversational turn taking. Along these lines, early non-verbal interactions establish the protocol by which children come to use language with others. And as they acquire language, they also come to integrate it into the preverbal turn-taking systems.

@@

The majority of work on children's early turn taking has focused on observations of spontaneous interaction. Children's first turn-like structures appear as early as two to three months in proto-conversation with their caregivers \citep{bruner1975, bruner1985}. During proto-conversations, caregivers interact with their infants as if they were capable of making meaningful contributions; they take every look, vocalization, arm flail and burp as ``utterances'' in the joint discourse \citep{bateson1975, jaffe2001, snow1977}. Infants catch onto the structure of proto-conversations quickly. By three to four months they notice disturbances to the contingency of their caregivers' response and, in reaction, change the rate and quality of their vocalizations \citep{k-bloom1988, masataka1993}. Infants at this age also notice changes to social contingency outside of turn structure. In the Still Face paradigm, caregivers interact with their infants and then suddenly halt, taking on a neutral expression with a sustained gaze. When faced with this sudden disappearance of social contingency, infants three months and older try a range of methods to reinitiate the interaction, such as vocalization, reaching, and smiling before looking away or getting upset \citep{rochat1998, toda1993}.

@@

We recruited 74 children between ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment. Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school. All participants were native English speakers. Approximately one-third (N=25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from analyses. None of the adult participants reported fluency in a second language.

@@

All participants (child and adult) completed a five-point calibration routine before the first video started. We used a dancing Elmo for the children's calibration image. During the experiment, participants watched all six 30-second conversation videos with 15--30 second filler videos in-between them (e.g., running puppies, singing muppets, flying bugs). The first and last conversations were in American English and the intervening conversations were Hebrew, Japanese, German, and Korean. The presentation order of the non-English videos was shuffled into four lists, which participants were assigned to randomly. The entire experiment, including instructions, took 10--15 minutes.

@@

As in the first experiment, we recorded participants' eye movements as they watched six short videos of dyadic conversation, and then analyzed their anticipatory glances from the current speaker to the upcoming speaker at points of turn transition.

@@

\subsubsection{Participants}
We recruited 27 undergraduate adults and 129 children between ages 1;0--6;11 to participate in our experiment. We recruited our child participants from the Children's Discovery Museum in San Jose, California, targeting approximately 20 children for each of the six 1-year age groups (range$=$20--23). All participants were native English speakers, though some parents (N$=$27) reported that their child heard a second (and sometimes third) language at home. None of the adult participants reported fluency in a second language. We ran Experiment 2 at a local children's museum because it gave us access to children with a more diverse range of ages.

@@

\subsection{Procedure}
We used the same experimental apparatus and procedure as in the first experiment. Each participant watched six puppet videos in random order, with five 15--30 second filler videos placed in-between (e.g., running puppies, moving balls, flying bugs). Three of the puppet videos had \textit{normal} audio while the other three had \textit{words only}, \textit{prosody only}, and \textit{no speech} audio. This experiment required no special instructions so the experimenter immediately began each session with calibration (same as before) and then stimulus presentation. The entire experiment took less than five minutes.

@@

Finally, we confirmed that young children showed a benefit for prosody over lexicosyntax. We conducted two \textit{post-hoc} \textit{t}-tests for children under three, comparing their anticipatory switching rate against their estimated random switching rate in the \textit{prosody only} and \textit{words only} conditions. We found that anticipatory gazes significantly exceeded chance for the \textit{prosody only} condition (\textit{t}(187.677)=-3.004, \textit{p}=.003), but not for the \textit{words only} condition (\textit{t}(229.828)=-0.8307, \textit{p}=.407). This finding suggests that there is indeed an early advantage for prosody over lexicosyntax in children under three. The advantage for prosody was not driven by a question effect; children and adults did not consistently show more switches after questions than non-questions in the \textit{prosody only} condition (Figure \ref{fig:questionsEN}). So, whereas the early advantage for prosody appeared for both transition types (question and non-question), the later-emerging advantage for lexicosyntax and prosody together was primarily limited to question transitions. This pattern of results may suggest a qualitative shift in the kind of predictions children make while observing a conversation, with linguistic cues (like prosody) playing different roles at different ages.

@@

\section{General Discussion}
\label{sec:gendisc}
Children begin to develop conversational turn-taking skills long before their first words \citep{bateson1975, hilbrinkInPrep, jaffe2001, snow1977}. As they acquire language, they also acquire the information needed to make accurate predictions about upcoming turn structure. Until recently, we have had very little data on how children weave language into their already-existing turn-taking behaviors.

@@

Our findings are only partially consistent with prior experimental work on turn taking. De Ruiter and colleagues (2006) found that adults' accuracy in identifying upcoming turn-ends was not significantly affected when intonation contours were flattened. Keitel and colleagues (2013) found the same for adults' anticipatory gaze switches, only finding that intonation flattening affected children's switches at 36 months (with switches by younger children never exceeding chance). Our results, in contrast, suggest that children's anticipatory switches emerge much earlier, and rely on prosody before they do lexicosyntax. Furthermore, our adult control participants patterned similarly to the children: They showed question effects for languages they didn't speak (Experiment 1) and showed the strongest question effects for lexicosyntax and prosody together (Experiment 2). Thus, neither the child nor the adult data support the idea that lexicosyntactic information is equal to full linguistic information in making predictions about upcoming turn structure \citep{de-ruiter2006}, at least in participants' spontaneous looking behavior. This may partially be due to the fact that we controlled syllable duration in addition to intonation, thereby eliminating two of the most salient cues to prosodic structure.

@@

Moving away from button press measures, Keitel and colleagues (2013) used observer gaze to investigate children's turn predictions, finding: (a) no evidence that children under three make switches above chance, (b) that three-year-olds' switches are significantly affected by flattened intonation, and (c) that adults' switches are not. Our current results are consistent with the general finding from their study, that prosody matters for young children's turn predictions. But our results also differ from theirs in a few important ways.

First, we found that children begin making above-chance anticipatory switches much earlier than three---our one-year-olds were already looking above chance in two of our four linguistic conditions in Experiment 2. This earlier emergence of anticipatory switching could be due to our more child-friendly stimuli, which we designed to be engaging and to mimic the kind of speech children often experience interactionally. Second, because we included more linguistic conditions to isolate the effects of prosody and lexicosyntax, we gained a broader perspective on how linguistic information changed children's predictions. Ultimately, the added conditions allowed us to find an early advantage for prosody and a later advantage for lexicosyntax.
